---
title: "Auto Process NY Covid-19 Case Data"
author: "John Minter"
date: "Started 2020-03-27, Last modified: 2020-03-29"
output:
  html_document:
    css: ../theme/jm-gray-vignette.css
    number_sections: yes
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA)
library(dplyr)
library(tibble)
library(here)
library(scales)
library(ggplot2)
library(RCurl)
library(kableExtra)
```

Set up the analysis. Use the `bVerbose` variable as a flag to decide whether
or not to print diagnostic data. The default is `FALSE`.

```{r printWorkDir}
bVerbose <- FALSE

if(bVerbose){
   print(getwd())
}

```

Next, load the data file and extract what we need using functions from the
`dplyr` package to create a `tibble` of values (an enhanced R dataframe that
works nicely with the `tidyverse` collection of R packages by Hadley Wickham.)
We will use the `kable` function from the `knitr` package to get a nice looking
table. We really only want the last few values...

```{r getNewYorkData}
df <- read.csv("./us-states.csv", header = TRUE, sep = ",")
df$date <- format(as.Date(df$date), "%m-%d")
tib <- as_tibble(df)
tib <- tib %>% filter(state == "New York")
tib  %>% select(date, cases, deaths) -> new_york

if(bVerbose){
   print(length(new_york$date))
   print(typeof(new_york$date[1]))
}
new_york$date <- as.Date(tib$date, "%m-%d")

if(bVerbose){
   head(new_york)
}

kable(tail(new_york))

if(bVerbose){
   print(typeof(tib$date))
   print(tib$date)
}
```

Next, we compute the death rate. We really only need the last value in the
table.

```{r computeDeathRate}

tot_deaths <- max(tib$deaths)
tot_cases <- max(tib$cases)

death_rate_pct <- 100*tot_deaths/tot_cases

death_rate_pct <- round(death_rate_pct, digits = 2)
```

New York's COVID-19 is `r death_rate_pct` percent.

Next, we plot the curve.

```{r plotNewYork}
plt <- ggplot(new_york, aes(x=date, y=cases)) +
       geom_line(colour='darkred', size=2) +
       xlab("Date") +
       ylab("Total Cases") + 
       ggtitle("NY COVID-19 Cases") +
       labs(caption = 'Data from Johns Hopkins') +
       scale_x_date() +
       # theme_minimal() +
       theme(axis.text=element_text(size=8),
            axis.title=element_text(size=12),
            plot.title=element_text(hjust = 0.5)) +
       NULL

print(plt)
```

Sadly, the curve has not flattened yet. Recall that this is for all of
New York State and is **dominated by cases in New York City**.

We will save the plot as in both the `png` and `jpg` formats.

```{r savePng, warning=FALSE, message=FALSE}
strOutPng <- "plt/Covid-19.png"

# we want the png to be close to 1024x768...
ggsave(plt, file=strOutPng, width=9.0, height=6.0,
       units="in", dpi=113.7778)

strOutJpg <- "plt/Covid-19.jpg"

# we want the png to be close to 1024x768...
ggsave(plt, file=strOutJpg, width=9.0, height=6.0,
       units="in", dpi=113.7778)
```
