---
title: "Auto Process Monroe County Covid-19 Case Data"
author: "John Minter"
date: "Started 2020-03-30, Last modified: 2020-04-05"
output:
  html_document:
    css: ../theme/jm-gray-vignette.css
    number_sections: yes
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA)
library(dplyr)
library(readr)
library(tibble)
library(here)
library(scales)
library(ggplot2)
library(RCurl)
library(kableExtra)
```

Set up the analysis. Use the `bVerbose` variable as a flag to decide whether
or not to print diagnostic data. The default is `FALSE`.

```{r printWorkDir}
bVerbose <- FALSE

if(bVerbose){
   print(getwd())
}

```

```{r pullData}
temp <- tempfile()
download.file("https://github.com/nytimes/covid-19-data/archive/master.zip",temp)
us_counties <- read_csv(unz(temp, "covid-19-data-master/us-counties.csv"))
path <- paste0(here(), "/us-counties.csv")
write_csv(us_counties, path, append = FALSE, col_names=TRUE)
us_states <- read_csv(unz(temp, "covid-19-data-master/us-states.csv"))
pa_states <- paste0(here(), "/us-states.csv")
write_csv(us_states, pa_states, append = FALSE, col_names=TRUE)
unlink(temp)
tail(us_counties)
tail(us_states)

```

Next, load the data file and extract what we need using functions from the
`dplyr` package to create a `tibble` of values (an enhanced R dataframe that
works nicely with the `tidyverse` collection of R packages by Hadley Wickham.)
We will use the `kable` function from the `knitr` package to get a nice looking
table. We really only want the last few values...

```{r getMonroeCountyData}
pa_cty <- paste0(here(), "/us-counties.csv")
df <- read.csv(pa_cty, header = TRUE, sep = ",")
df$date <- format(as.Date(df$date), "%m-%d")
tib <- as_tibble(df)
tib$date <- as.Date(df$date, "%m-%d")
ny <- tib %>% filter(state == "New York")
monroe_cty <- ny %>% filter(county == "Monroe")
monroe_cty  %>% select(date, cases, deaths) -> mc_vals
tail(mc_vals)
```

```{r tailMc}
tail(mc_vals)
```



```{r listCases}
print(mc_vals)
```

Next, we compute the total deaths for Monroe County

```{r computeDeathRate}

tot_deaths <- sum(monroe_cty$deaths)
tot_cases <- sum(monroe_cty$cases)

death_rate_pct <- 100*tot_deaths/tot_cases

death_rate_pct <- round(death_rate_pct, digits = 2)
```


Monroe County's COVID-19 has `r tot_cases` cases and `r tot_deaths` deaths
with a death rate of `r death_rate_pct` percent.

Next, we plot the curve.

```{r plotMonroeCounty}
mc_tot_plt <- ggplot(mc_vals, aes(x=date, y=cases)) +
              geom_line(colour='darkred', size=2) +
              xlab("Date") +
              ylab("Total Cases") + 
              ggtitle("Monroe County COVID-19 Cases") +
              labs(caption = 'Data from Johns Hopkins') +
              scale_x_date() +
              theme(axis.text=element_text(size=8),
                    axis.title=element_text(size=12),
                    plot.title=element_text(hjust = 0.5)) +
              NULL

print(mc_tot_plt)
```


It looks like Monroe County's curve is starting to flatten.

We will save the plot as in both the `png` and `jpg` formats.

```{r savePng, warning=FALSE, message=FALSE}
strOutPng <- "plt/Monroe_Cty_Covid-19.png"

# we want the png to be close to 1024x768...
ggsave(mc_tot_plt, file=strOutPng, width=9.0, height=6.0,
       units="in", dpi=113.7778)

strOutJpg <- "plt/Monroe_Cty_Covid-19.jpg"

# we want the png to be close to 1024x768...
ggsave(mc_tot_plt, file=strOutJpg, width=9.0, height=6.0,
       units="in", dpi=113.7778)
```
