---
title: "Auto Process CDC Data"
author: "John Minter"
date: "Started 2020-04-05, Last modified: 2020-04-05"
output:
  html_document:
    css: ../theme/jm-gray-vignette.css
    number_sections: yes
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA)
library(dplyr)
library(readr)
library(tibble)
library(here)
library(scales)
library(ggplot2)
library(RCurl)
library(kableExtra)
```

# Plot 2019 Flu Data from CDC

Set up the analysis. Use the `bVerbose` variable as a flag to decide whether
or not to print diagnostic data. The default is `FALSE`.

```{r printWorkDir}
bVerbose <- FALSE

if(bVerbose){
   print(getwd())
}

```

# Load the CDC data



Next, load the data file and extract what we need using functions from the
`dplyr` package to create a `tibble` of values (an enhanced R dataframe that
works nicely with the `tidyverse` collection of R packages by Hadley Wickham.)
We will use the `kable` function from the `knitr` package to get a nice looking
table. We really only want the last few values...

```{r getFluData, warning=FALSE}
pa_flu <- paste0(here(), "/R/CDC-flu-fixed.csv")
# pa_flu
flu <- read.csv(pa_flu, header = TRUE, sep = ",")
flu$Date_Meas=as.Date(flu$Date_Meas)
flu$A_Cum_Sum <- cumsum(flu$Total_A)
flu$B_Cum_Sum <- cumsum(flu$Total_B)
head(flu)
```

```{r printTable, warning=FALSE}
kable(tail(flu))
```


Next, we plot Flu A.

```{r plotFlu}
plt_flu <- ggplot(flu) +
           geom_line(data=flu, aes(x=Date_Meas, y=Total_A),
                      colour="darkblue") +
           geom_line(data=flu, aes(x=Date_Meas, y=Total_B),
                      colour="darkred") +
           xlab("Week (2019-2020)") +
           ylab("Total Cases Flu") +
           ggtitle("Fall 2019 Flu Season") +
           labs(caption = 'Data from CDC') +
           scale_x_date(date_breaks = "1 week", date_labels = "%W") +
           # scale_x_date() +
           # theme_minimal() +
           theme(axis.text=element_text(size=8),
                 axis.title=element_text(size=12),
                 plot.title=element_text(hjust = 0.5)) +
           NULL

print(plt_flu)
  
```

We will save the plot as in both the `png` and `jpg` formats.

```{r savePngFlu, warning=FALSE, message=FALSE}
strOutPng <- "plt/Flu2019-2020-Cases.png"

# we want the png to be close to 1024x768...
ggsave(plt_flu, file=strOutPng, width=9.0, height=6.0,
       units="in", dpi=113.7778)

strOutJpg <- "plt/Flu2019-2020-Cases.jpg"

# we want the png to be close to 1024x768...
ggsave(plt_flu, file=strOutJpg, width=9.0, height=6.0,
       units="in", dpi=113.7778)
```



```{r plotFluCum}
plt_flu_cum <- ggplot(flu) +
               geom_line(data=flu, aes(x=Date_Meas, y=A_Cum_Sum),
                         colour="darkblue") +
               geom_line(data=flu, aes(x=Date_Meas, y=B_Cum_Sum),
                         colour="darkred") +
               xlab("Week (2019-2020)") +
               ylab("Cumulative Cases Flu") +
               ggtitle("Fall 2019 Flu Season") +
               labs(caption = 'Data from CDC') +
               scale_x_date(date_breaks = "1 week", date_labels = "%W") +
               # scale_x_date() +
              # theme_minimal() +
              theme(axis.text=element_text(size=8),
                    axis.title=element_text(size=12),
                    plot.title=element_text(hjust = 0.5)) +
              NULL

print(plt_flu_cum)


```

We will save the plot as in both the `png` and `jpg` formats.

```{r savePngFlu2, warning=FALSE, message=FALSE}
strOutPng <- "plt/Flu2019-2020-CumCases.png"

# we want the png to be close to 1024x768...
ggsave(plt_flu_cum, file=strOutPng, width=9.0, height=6.0,
       units="in", dpi=113.7778)

strOutJpg <- "plt/Flu2019-2020-CumCases.jpg"

# we want the png to be close to 1024x768...
ggsave(plt_flu_cum, file=strOutJpg, width=9.0, height=6.0,
       units="in", dpi=113.7778)
```
