---
title: "Auto Process New York and Monroe County Covid-19 Case Data"
author: "John Minter"
date: 'Started 2020-03-27, Last modified: 2020-06-10'
output:
  html_document:
    css: ../theme/jm-gray-vignette.css
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA)
library(dplyr)
library(readr)
library(tibble)
library(here)
library(scales)
library(ggplot2)
library(RCurl)
library(kableExtra)
```

# Plot Total COVID-19 Cases

Set up the analysis. Use the `bVerbose` variable as a flag to decide whether
or not to print diagnostic data. The default is `FALSE`.

```{r setupFlags}
bVerbose <- FALSE
bDownload <- FALSE

if(bVerbose){
   print(getwd())
}

```

Pull down the data


```{r pullData}
if(bDownload){
  temp <- tempfile()
  download.file("https://github.com/nytimes/covid-19-data/archive/master.zip",temp)
  us_counties <- read_csv(unz(temp, "covid-19-data-master/us-counties.csv"))
  path <- paste0(here(), "/us-counties.csv")
  write_csv(us_counties, path, append = FALSE, col_names=TRUE)
  us_states <- read_csv(unz(temp, "covid-19-data-master/us-states.csv"))
  path <- paste0(here(), "/us-states.csv")
  write_csv(us_states, path, append = FALSE, col_names=TRUE)
  unlink(temp)
  tail(us_counties)
  tail(us_states)
}

```

Next, load the data file and extract what we need using functions from the
`dplyr` package to create a `tibble` of values (an enhanced R dataframe that
works nicely with the `tidyverse` collection of R packages by Hadley Wickham.)
We will use the `kable` function from the `knitr` package to get a nice looking
table. We really only want the last few values...

```{r getNewYorkData}
pa_states <- paste0(here(), "/us-states.csv")
df <- read.csv(pa_states, header = TRUE, sep = ",")
df$date <- format(as.Date(df$date), "%m-%d")
tib <- as_tibble(df)
tib <- tib %>% filter(state == "New York")
tib  %>% select(date, cases, deaths) -> new_york

if(bVerbose){
   print(length(new_york$date))
   print(typeof(new_york$date[1]))
}
new_york$date <- as.Date(tib$date, "%m-%d")

if(bVerbose){
   head(new_york)
}

kable(tail(new_york))

if(bVerbose){
   print(typeof(tib$date))
   print(tib$date)
}
```

Next, we compute the death rate. We really only need the last value in the
table.

```{r computeDeathRate}

tot_deaths <- max(tib$deaths)
tot_cases <- max(tib$cases)

death_rate_pct <- 100*tot_deaths/tot_cases

death_rate_pct <- round(death_rate_pct, digits = 2)
```

New York's COVID-19 death rate is `r death_rate_pct` percent.

Next, we plot the curve.

```{r plotNewYorkTotal}
plt_ny_tot <- ggplot(new_york, aes(x=date, y=cases)) +
              geom_line(colour='darkred', size=2) +
              xlab("Date") +
              ylab("Total Cases") + 
              ggtitle("NY Total COVID-19 Cases") +
              labs(caption = 'Data from New York Times') +
              scale_x_date() +
              # theme_minimal() +
              theme(axis.text=element_text(size=8),  axis.title=element_text(size=12), plot.title=element_text(hjust = 0.5)) +
              NULL

print(plt_ny_tot)
```

The curve is starting to flatten. Recall that this is for all of
New York State and is **dominated by cases in New York City**.


We will save the plot as in both the `png` and `jpg` formats.

```{r savePng, warning=FALSE, message=FALSE}
strOutPng <- "plt/Covid-19-Total-Cases-NY.png"

# we want the png to be close to 1024x768...
ggsave(plt_ny_tot, file=strOutPng, width=9.0, height=6.0,
       units="in", dpi=113.7778)

strOutJpg <- "plt/Covid-19-Total-Cases-NY.jpg"

# we want the png to be close to 1024x768...
ggsave(plt_ny_tot, file=strOutJpg, width=9.0, height=6.0,
       units="in", dpi=113.7778)
```

# Plot NY COVID Deaths

```{r plotNewYorkDeaths}
plt_ny_deaths <- ggplot(new_york, aes(x=date, y=deaths)) +
                 geom_line(colour='darkred', size=2) +
                 xlab("Date") +
                 ylab("Total NY Deaths") + 
                 ggtitle("NY Total COVID-19 Deaths") +
                 labs(caption = 'Data from the New York Times') +
                 scale_x_date() +
                 # theme_minimal() +
                 theme(axis.text=element_text(size=8),
                 axis.title=element_text(size=12),
                 plot.title=element_text(hjust = 0.5)) +
                 NULL

print(plt_ny_deaths)
```

We will save the plot as in both the `png` and `jpg` formats.

```{r savePngDeaths, warning=FALSE, message=FALSE}
strOutPng <- "plt/Covid-19-Total-Deaths-NY.png"

# we want the png to be close to 1024x768...
ggsave(plt_ny_deaths, file=strOutPng, width=9.0, height=6.0,
       units="in", dpi=113.7778)

strOutJpg <- "plt/Covid-19-Total-Deaths-NY.jpg"

# we want the png to be close to 1024x768...
ggsave(plt_ny_deaths, file=strOutJpg, width=9.0, height=6.0,
       units="in", dpi=113.7778)
```

# Plot COVID New York New Cases


Make a tibble of new cases

```{r testDifference}
new_cases <- diff(tib$cases)
the_date <- as.Date(tib$date[-1], "%m-%d")
print(class(the_date))

new_cases_tbl <- data.frame(the_date, new_cases)
print(head(new_cases_tbl))
print(tail(new_cases_tbl))
```

```{r plotNewYorkTotalNewCases}
plt_ny_new_cases <- ggplot(new_cases_tbl, aes(x=the_date, y=new_cases)) +
                    geom_line(colour='darkred', size=2) +
                    xlab("Date") +
                    ylab("New Cases") + 
                    ggtitle("NY Total New COVID-19 Cases") +
                    labs(caption = 'Data from New York Times') +
                    scale_x_date() +
                    theme(axis.text=element_text(size=8),
                    axis.title=element_text(size=12),
                    plot.title=element_text(hjust = 0.5)) +
                    NULL

print(plt_ny_new_cases)

```

Save the New Cases Plot

```{r savePngNew, warning=FALSE, message=FALSE}
strOutPng <- "plt/Covid-19-New-Cases-NY.png"

# we want the png to be close to 1024x768...
ggsave(plt_ny_new_cases, file=strOutPng, width=9.0, height=6.0,
       units="in", dpi=113.7778)

strOutJpg <- "plt/Covid-19-New-Cases-NY.jpg"

# we want the png to be close to 1024x768...
ggsave(plt_ny_new_cases, file=strOutJpg, width=9.0, height=6.0,
       units="in", dpi=113.7778)
```


# Process Monroe County Data

Next, load the data file and extract what we need using functions from the
`dplyr` package to create a `tibble` of values (an enhanced R dataframe that
works nicely with the `tidyverse` collection of R packages by Hadley Wickham.)
We will use the `kable` function from the `knitr` package to get a nice looking
table. We really only want the last few values...

```{r getMonroeCountyData}
pa_cty <- paste0(here(), "/us-counties.csv")
df <- read.csv(pa_cty, header = TRUE, sep = ",")
df$date <- format(as.Date(df$date), "%m-%d")
tib <- as_tibble(df)
tib$date <- as.Date(df$date, "%m-%d")
ny <- tib %>% filter(state == "New York")
monroe_cty <- ny %>% filter(county == "Monroe")
monroe_cty  %>% select(date, cases, deaths) -> mc_vals
tail(mc_vals)
```
Next, we compute the total deaths for Monroe County

```{r computeMCDeathRate}

tot_deaths <- max(monroe_cty$deaths)
tot_cases <- max(monroe_cty$cases)

death_rate_pct <- 100*tot_deaths/tot_cases

death_rate_pct <- round(death_rate_pct, digits = 2)
```


Monroe County's COVID-19 has `r tot_cases` cases and `r tot_deaths` deaths
with a death rate of `r death_rate_pct` percent.

Next, we plot the curve.

```{r plotMonroeCounty}
mc_tot_plt <- ggplot(mc_vals, aes(x=date, y=cases)) +
              geom_line(colour='darkred', size=2) +
              xlab("Date") +
              ylab("Total Cases") + 
              ggtitle("Monroe County COVID-19 Cases") +
              labs(caption = 'Data from New York Times') +
              scale_x_date() +
              theme(axis.text=element_text(size=8),
                    axis.title=element_text(size=12),
                    plot.title=element_text(hjust = 0.5)) +
              NULL

print(mc_tot_plt)
```

We will save the plot as in both the `png` and `jpg` formats.

```{r saveMCPng, warning=FALSE, message=FALSE}
strOutPng <- "plt/Monroe_Cty_Covid-19.png"

# we want the png to be close to 1024x768...
ggsave(mc_tot_plt, file=strOutPng, width=9.0, height=6.0,
       units="in", dpi=113.7778)

strOutJpg <- "plt/Monroe_Cty_Covid-19.jpg"

# we want the png to be close to 1024x768...
ggsave(mc_tot_plt, file=strOutJpg, width=9.0, height=6.0,
       units="in", dpi=113.7778)
```

## Reproducibility

<details><summary>Reproducibility receipt</summary>

```{r}
## datetime
Sys.time()

## repository
git2r::repository()

## session info
sessionInfo()
```

</details>